---
title: "Plot assignment"
author: "Template: Ian Hussey; content: Isabel Mayer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(psych)
library(patchwork)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(see)
library(ggsignif)

```

# Data

```{r}
mres <- readr::read_csv("data/RMET_Project_Data_OSF.csv")
```

## Removing straight liners
Taken over from the original processing code.
```{r}
#REMOVING STRAIGHT LINERS - SKIP THIS SECTION TO ANALYSE DATA WITH THESE PARTICIPANTS RETAINED


# # identifying participants who straight line responded to the AQ

a <- dplyr::select(mres,starts_with("AQ")) 
a<-na.omit(a)

keep.a <- apply(a,1,function(x) length(unique(x[!is.na(x)])) !=1) 
retained_a<-a[keep.a,]
excluded_a <-a[!keep.a,]
which(!keep.a)

# identifying participants who straight line responded to TAS
t <- dplyr::select(mres,starts_with("TAS"))
keep.t <- apply(t,1,function(x) length(unique(x[!is.na(x)])) !=1)
retained_t<-t[keep.t,]
which(!keep.t)

# # identifying participants who straight line responded to IMT
tom <- mres[,c(paste0("TOM0",1:9),paste0("TOM",10:16))]
keep.tom <- apply(tom,1,function(x) length(unique(x[!is.na(x)])) !=1)
retained_tom<-t[keep.tom,]
which(!keep.tom)


# removing straight line responing participants                                                                                
mres.pass <- mres[-c(20 ,  28,  100,  152, 376 , 549 , 665, 670,  737 , 749 ,782,794, 825,832,847, 860,  864,  871 , 903 ,904,920, 927,936,960,975,  977, 980,1004,1007,1023,1059, 1061, 1086,1089, 1091 ,1107,1161, 1163,1210,1214,1221  ),]                                                                                               

mres <- mres.pass # renaming as mres
mres <- readr::type_convert(mres) # converting the types of the data
mres.tom <- mres %>% drop_na  (TOM16) #remove participants who did not complete ToM task.

```

## Reverse Code
Taken from original processing code.
```{r}
#CODE TO REVERSE CODE TAS

mres$TAS04 <- 6 - mres$TAS04
mres$TAS05 <- 6 - mres$TAS05
mres$TAS10 <- 6 - mres$TAS10
mres$TAS18 <- 6 - mres$TAS18
mres$TAS19 <- 6 - mres$TAS19

# CODE TO REVERSE CODE AQ 

mres$AQ02 <- 5 - mres$AQ02
mres$AQ04 <- 5 - mres$AQ04
mres$AQ05 <- 5 - mres$AQ05
mres$AQ07 <- 5 - mres$AQ07
mres$AQ10 <- 5 - mres$AQ10
mres$AQ13 <- 5 - mres$AQ13
mres$AQ14 <- 5 - mres$AQ14
mres$AQ15 <- 5 - mres$AQ15
mres$AQ16 <- 5 - mres$AQ16
mres$AQ22 <- 5 - mres$AQ22
mres$AQ23 <- 5 - mres$AQ23
mres$AQ25 <- 5 - mres$AQ25
mres$AQ26 <- 5 - mres$AQ26

#CODE TO SCORE IMT FALSE ANSWERS

mres$TOM02 <- 3 - mres$TOM02
mres$TOM04 <- 3 - mres$TOM04
mres$TOM05 <- 3 - mres$TOM05
mres$TOM07 <- 3 - mres$TOM07
mres$TOM11 <- 3 - mres$TOM11
mres$TOM13 <- 3 - mres$TOM13
mres$TOM15 <- 3 - mres$TOM15
mres$TOM16 <- 3 - mres$TOM16
```

## Totals
Taken from original processing code.
```{r}
# CREATING TOTALS DATAFRAME

# LIST OF TAS TOTALS OVERALL AND FOR SUBSCALES 
tas.list <- list(tas.total=c("TAS01", "TAS02", "TAS03", "TAS04", "TAS05", "TAS06", "TAS07", "TAS08", "TAS09", "TAS10", "TAS11", "TAS12", "TAS13","TAS14", "TAS15", "TAS16", "TAS17", "TAS18", "TAS19", "TAS20"), 
                 tas.identify=c("TAS01", "TAS03", "TAS06", "TAS07", "TAS09","TAS13", "TAS14"),
                 tas.describe=c("TAS02","TAS04", "TAS11","TAS12","TAS17"),
                 tas.external=c("TAS05","TAS08","TAS10","TAS15","TAS16","TAS18", "TAS19","TAS20"))


tas.metrics <- scoreItems(tas.list, mres, totals=TRUE) #psychometric data for TAS results
tas.scores <- as.data.frame(tas.metrics$scores) #TAS scores with total and subscales
tas.scores$ID.2 <- mres$ID #adding ID column

# LIST OF AQ TOTALS OVERALL AND SUBSCALES 

aq.list <- list(aq.total=c("AQ01", "AQ02", "AQ03", "AQ04", "AQ05", "AQ06", "AQ07", "AQ08", "AQ09", "AQ10", "AQ11", "AQ12", "AQ13", "AQ14","AQ15","AQ16", "AQ17", "AQ18", "AQ19", "AQ20", "AQ21", "AQ22", "AQ23", "AQ24", "AQ25", "AQ26", "AQ27", "AQ28"),
                aq.social.behaviour=c("AQ01", "AQ02", "AQ03", "AQ04",  "AQ06", "AQ08", "AQ09", "AQ10", "AQ11", "AQ12",  "AQ14", "AQ15", "AQ17", "AQ18", "AQ19", "AQ20", "AQ21", "AQ23", "AQ24", "AQ25", "AQ26", "AQ27", "AQ28"),
                aq.numbers.patterns=c("AQ05","AQ07","AQ13","AQ16","AQ22"),
                aq.social.skills=c("AQ01","AQ09","AQ10","AQ12","AQ15","AQ24","AQ26","AQ27"),
                aq.routine=c("AQ02","AQ17","AQ19","AQ26"),
                aq.switching=c("AQ04","AQ08","AQ18","AQ21"),
                aq.imagination=c("AQ03","AQ06","AQ11","AQ14","AQ20","AQ23","AQ25","AQ28"))

aq.metrics <- scoreItems(aq.list, mres, totals=TRUE) #psychometric data for AQ results
aq.scores <- as.data.frame(aq.metrics$scores) #AQ scores with total and subscale totals
aq.scores$ID.1 <- mres$ID #Add participant ID column 

#RMET SCORES

rmet.scores <- mres[,c(paste0("R0",1:9),paste0("R",10:36))] #creating table of RMET responses
# run exploratory code for items that fail consensus here to keep all four choices separate.
rmet.scores[rmet.scores !=1] <- 0
rmet.scores$rmet.correct <- rowSums(rmet.scores==1) #add a column to rmet.scores with number of RMET correct responses, which equals RMET score.
rmet.scores$ID <- mres$ID #add ID column to rmet.scores
rmet.scores$aq.total <- aq.scores$aq.total #ad aq total scores column to rmet.scores
rmet.scores <- dplyr::select(rmet.scores,ID,aq.total,rmet.correct,everything()) #reordering columns

# IMT TOM and MEMORY SCORES
imt.scores <- dplyr::select(mres, TOM01, TOM02, TOM03, TOM04, TOM05, TOM06, TOM07, TOM08,  TOM09, TOM10, TOM11, TOM12, TOM13, TOM14, TOM15, TOM16)
imt.scores$ID <- mres$ID # add ID column
imt.scores$imt.correct <- rowSums(imt.scores ==1) #total score for all IMT questions
imt.scores <- dplyr::select(imt.scores, ID, imt.correct, everything()) #reorder to have ID and total score columns first

imt.memory.scores <- dplyr::select(imt.scores,ID, TOM03, TOM04,TOM06, TOM07, TOM10, TOM12, TOM15, TOM16 ) # IMT memory items
imt.tom.scores <- dplyr::select(imt.scores, ID, TOM01,TOM02,TOM05,TOM08, TOM09, TOM11, TOM13, TOM14) #IMT ToM items
imt.memory.scores$imt.memory.correct <- rowSums(imt.memory.scores==1) #total score for IMT memory
imt.memory.scores <- dplyr::select(imt.memory.scores, ID, imt.memory.correct, everything()) #reorder columns
imt.tom.scores$imt.tom.correct <- rowSums(imt.tom.scores==1) # total score for IMT ToM
imt.tom.scores <- dplyr::select(imt.tom.scores, ID, imt.tom.correct, everything())

# creating a data frame with totals for all meausres
totals <- tibble(ID=mres$ID, attention=mres$Attention_1, ethnicity=mres$ethnicity, gender=mres$gender, age=mres$age, god=mres$Dem_god_4, comfort=mres$R_Comfort_1, tas.scores, aq.scores, imt.correct=imt.scores$imt.correct, imt.memory.correct=imt.memory.scores$imt.memory.correct,imt.tom.correct= imt.tom.scores$imt.tom.correct, rmet.correct=rmet.scores$rmet.correct, education=mres$Dem_education) #totals of scores for measures
```
Note: I added the variable mres$ethnicity to the totals dataframe for plotting purposes.

# Good Plot
Plot showing the distribution of Scores in Autism Spectrum Quotient Short (AQ-S), Toronto Alexithymia Scale (TAS) and Reading the Mind in the Eyes test (RMET)
```{r}
# Add TAS variable
totals <- totals |>
  mutate(TAS = if_else(tas.total >54, "Alexithymia", "No Alexithymia"))

totals$TAS <- factor(totals$TAS)

plot_good <- ggscatter(totals, x = "rmet.correct", y = "aq.total", 
                          palette = "jco", facet.by = "TAS",
                          add = "reg.line") + geom_jitter(position = position_jitter(width = 0.05), alpha = 0.5, size = 0.005, shape = 16, show.legend = FALSE) 

# Add labels
plot_good <- plot_good +
  labs(
    x = "RMET Score",
    y = "AQ-S Score",
    title = "Autism Spectrum Quotient and Reading the Mind in the Eye Test Score",
    subtitle = "For people with high and low scores on the Toronto Alexithymia Scale",
    caption = "Source: https://osf.io/6hy95"
  )
print(plot_good)
```


Reasons why this is a good plot:
- The scatter plot allows to visualize the relationship between the Autism Spectrum Quotient (AQ) score (x = "aq.total") and the Reading the Mind in the Eyes test (RMET) score (y = "rmet.correct"). Putting these variables in comparison makes sense, since one can expect a relationship.
- In addition, the plot is faceted by the TAS Score (I set the Cut-Off 54 for Alexithymia), allowing us to compare the scatter plots side by side. 
- The addition of jitter helps avoid overplotting, making it easier to see the density of points.
- The regression lines help to visualize the trend of the correlation at first sight.
- Proper labeling and legends make the plot more informative. 
- The title is long but informative
- The source is added at the bottom right hand corner
- There is no color added since it is not necessary in this context (the alexithymia/no alexithymia distinction is clear because of the facets) - however one could add colours for aesthetics:)
- 

# Bad Plot
Plot showing the relationship of ethnicity, the score in the Imposing Memory Task (IMT) and the level of reported comfort when viewing eye stimuli
```{r}
plot_bad <- ggplot(totals, aes(x = imt.correct, y = comfort, color = ethnicity)) +
  geom_point(size = 5, alpha = 0.5) + 
  labs(title = "Bad Plot with Inappropriate Choices") +
  geom_text(aes(label = ID), vjust = 1, hjust = 0.75) +
  theme_minimal()  


# Print the bad plot
print(plot_bad)

```


Reasons why this is a bad plot:
- A scatter plot (geom_point()) is used, but it does not convey any meaningful information. The plot shows the relationship between the variables imt.correct (score of the IMT) and comfort (reported level of comfort when being presented with eye stimuli for the RMET) despite there beeing no reason to assume a relationship. The density of the jitter points is unclear because of the coloring.  
- There are no axis labels (xlab() and ylab()) or a legend (labs(color = ...)) to provide context for the variables. The axis therefore carry the name of the column in the dataset (i.e. imt.correct) which doesn't look nice and is confusing to the viewer. 
- Color is mapped to the "ethnicity" variable, but the coding in numbers (1-16) remains unexplained, the colors do not convey any meaningful information and they overlap in the plot. 
- The jitter points are unnecessarily labelled with the ID numbers but they are not legible as they too overlap. They are set slightly off to the left which looks unsettling and doesn't aid the legibility.
- The title doesn't convey any meaningful information.

# Write to disk

```{r}
ggsave("plot_good_IsabelMayer.png", plot = plot_good, width = 8, height = 5)
ggsave("plot_bad_IsabelMayer.png", plot = plot_bad, width = 8, height = 6)
```

